---
title: "pelodiscussinensisPH"
author: "Axela Keeza T. Ugalde"
date: "`r Sys.Date()`"
output: html_document
---
# Activity 10 - Geographic Information System: Creating Your First Map {.tabset .tabset-fade .tabset-pills}
  
## A. Retrieving spatial data
```{r setup, echo = TRUE, warning = TRUE, fig.align = "center"}
library(sf)
library(terra)
library(dplyr)
library(ggplot2)
library(ggnewscale)
library(tidyverse)

# Read CSV
my_data <- read.csv("occurrence - Pelodiscus sinensis.csv")

# Clean and Subset
clean_data <- my_data |>
  filter(!is.na(decimalLatitude) & !is.na(decimalLongitude))

target_provinces <- st_read("PHL_adm1.shp") |>
  st_transform(4326) |>
  st_make_valid()

# Elevation
phil_dem <- rast("ph_ strm_geo.tif")
dem_resampled <- aggregate(phil_dem, fact = 4, fun = mean)

if (st_crs(target_provinces) != st_crs(dem_resampled)) {
  target_provinces <- st_transform(target_provinces, st_crs(dem_resampled))
}
dem_clipped <- crop(dem_resampled, target_provinces) |> mask(target_provinces)

dem_df <- as.data.frame(dem_clipped, xy = TRUE)
colnames(dem_df) <- c("x", "y", "elevation")

plot(dem_clipped)
```


## B. Map
```{r map, echo = TRUE, warning = TRUE, fig.align = "center"}
ggplot() +
  geom_tile(data = dem_df, aes(x = x, y = y, fill = elevation)) +
 scale_fill_gradientn(
    colors = c("#1a4d2e", "#4f772d", "#90a955", "#ecf39e", "#f3c68f", "#ee6c4d", "#c32f27"),
    values = c(0, 0.05, 0.15, 0.30, 0.50, 0.75, 1.0), 
    name = "Elevation (m)"
  ) +

  geom_sf(data = target_provinces, fill = NA, 
          color = "gray30", 
          linewidth = 0.1) +
  
  geom_point(data = clean_data, aes(x = decimalLongitude, y = decimalLatitude), 
             shape = 21, 
             fill = "hotpink", 
             color = "black", 
             size = 2.0, 
             stroke = 0.5) +

  coord_sf() + 
  labs(title = expression(paste("National Distribution of ", italic("Pelodiscus sinensis"))),
       subtitle = "GBIF Data Philippines",
       x = "Longitude (°E)", y = "Latitude (°N)" +
  
  theme_linedraw() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    panel.background = element_rect(fill = "aliceblue"),
    panel.grid.major = element_line(color = "gray90", 
                                    linetype = "dashed", 
                                    linewidth = 0.3)
  ))


ggsave("philippines_distribution_map.pdf", 
       width = 8, 
       height = 11, 
       units = "in", 
       dpi = 300)